{RILS}-{ROLS} : Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-12-01 paper_resources/doc/bmc_jbd/bmc_jbd.tex
BMC template: converted

\documentclass[a4paper,12pt]{elsarticle}
% vim: tw=0 wm=0

\setcounter{tocdepth}{3}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{comment}
\usepackage{placeins}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{enumitem}
\usepackage[utf8]{inputenc}
\usepackage{booktabs}
\usepackage{array}
\usepackage[pdfencoding=auto,psdextra]{hyperref}
\usepackage{booktabs}
\usepackage{bookmark}% faster updated bookmarks
\usepackage{hypcap} % fix the links
\evensidemargin\oddsidemargin
\usepackage{graphicx}
\pagestyle{plain}
\usepackage{xcolor}
\newcommand\ToDo[1]{\textcolor{red}{#1}}
\usepackage{tabularx}
\usepackage{xspace}
\usepackage{color}
\usepackage{epsfig}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{mathrsfs}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
%\usepackage{changes}
\usepackage{tikz}
\usepackage{fullpage}
\usepackage{calc}
\usetikzlibrary{positioning,shadows,arrows,trees,shapes,fit}
\usepackage{blindtext}
\usepackage{pgfplots}
\pgfplotsset{compat=1.16}

\usepackage{numprint}
\npdecimalsign{.}
\npthousandsep{}
\usepackage{pythonhighlight}

\usepackage[draft,nomargin,inline]{fixme}
\fxsetface{inline}{\itshape}
\fxsetface{env}{\itshape}
%\fxuselayouts{margin}
%\fxuselayouts{inline}
\fxusetheme{color}

\usepackage{url}
\newcommand{\keywords}[1]{\par\aDSvspace\baselineskip
	\noindent\keywordname\enspace\ignorespaces#1}

\usepackage{tikz}
\usetikzlibrary{positioning}
\definecolor{canaryyellow}{rgb}{1.0, 0.94, 0.0}
\definecolor{brightgreen}{rgb}{0.4, 1.0, 0.0}
\definecolor{jazzberryjam}{rgb}{0.65, 0.04, 0.37}

%defining of command

\newcommand\floor[1]{\lfloor#1\rfloor}
\newcommand\ceil[1]{\lceil#1\rceil}
\newcommand\str[1]{\texttt{#1}}
\newcommand\pL[1][]{\ensuremath{p^{\mathrm{L}#1}}}
\newcommand\pR[1][]{\ensuremath{p^{\mathrm{R}#1}}}
\newcommand\qL{\ensuremath{q^\mathrm{L}}}
\newcommand\qR{\ensuremath{q^\mathrm{R}}}
\newcommand\pLH{\ensuremath{\hat{p}^\mathrm{L}}}
\newcommand\pRH{\ensuremath{\hat{p}^\mathrm{R}}}
\newcommand{\Vext}{\ensuremath{V_\mathrm{{ext}}}}
\newcommand\UB{\ensuremath{\mathrm{UB}}}
\newcommand\Sigmand{\ensuremath{\Sigma^\mathrm{nd}}}
\newcommand{\mdmwnpp}{MDMWNPP\xspace}

\renewcommand{\labelenumii}{\theenumii}
\renewcommand{\theenumii}{\theenumi.\arabic{enumii}.}
\setlength{\leftmarginii}{1.8ex}
\raggedbottom
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}

@@

\begin{document}

	%\setlength{\parindent}{0pt}  % disallow indentations
	%\numberwithin{table}{1}
	%\mainmatter  % start of an individual contribution

	% first the title is needed
	\title{\textsc{RILS}-\textsc{ROLS} : Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares}

	\author[1]{Aleksandar Kartelj}
	\author[2]{Marko Djukanovi\'c}
	\address[1]{$kartelj@matf.bg.ac.rs$, \\  Faculty of Mathematics, University of Belgrade, Serbia}
	\address[2]{$ marko.djukanovic@pmf.unibl.org$,\\   Faculty of Natural Sciences and Mathematics, University of Banja Luka, Bosnia and Herzegovina}
	\begin{abstract}
		In this paper, we solve the well-known symbolic regression problem. This problem has been intensively studied in the literature and has a wide range of applications. The symbolic regression models could provide insights towards establishing physical theory behind experimentally observed data. Practical applications include revealing complex ecological dynamics in ecology, discovering mutations effects on protein stability in biology, modeling analytic representations of the exciton binding energy in energy science, finding models for band gaps of NaCl-type compounds
		in material science, just to name a few.

@@

		The experiments are conducted on the two well-known ground-truth benchmark sets called \textsc{Feynman} and \textsc{Strogatz}. \textsc{RILS}-\textsc{ROLS} was compared to 14 other competitors from the literature with  presence of different levels of the Gaussian white noise in the input data. Our method outperformed all 14 competitors with respect to the true symbolic model accuracy percentage under all level of noise. 
		In addition to evaluation on known ground-truth datasets, a new randomly generated set of problem instances is introduced. The goal of \textsc{Random} dataset was to test the scalability of our method with respect to incremental equation sizes and different levels of noise. \textsc{RILS}-\textsc{ROLS} can be considered as a new state-of-the-art method for solving the problem of symbolic regression when ground-truth equations are known.
	\end{abstract}
	\maketitle

	\section{Introduction}\label{sec:introduction}

@@

	%\section{Literature review}\label{sec:lit-rev}

@@

	Finally, by setting each partial derivative to zero, we get the system of $d+1$ linear equations with $d+1$ unknowns ($\mathbf{k}=(k_1, ..., k_d)$ and $b$): 
	\begin{align*}
		& \sum_{i=1}^{n} -2x_{i1} (y_i - \tilde{y}_i) = 0 \\
		& \vdots \\
		& \sum_{i=1}^{n} -2x_{id} (y_i - \tilde{y}_i) = 0 \\
		&\sum_{i=1}^{n} -2  (y_i - \tilde{y}_i) = 0.
	\end{align*}
	%The above system can be transformed to 
	%\begin{align*}
	%	 & \sum y_i \cdot x_i   = b\sum x_i \cdot x_i  +  k \cdot \sum x_i \\
	%	 & \sum y_i  = k \sum x_i + b \cdot d. 
	%  \end{align*}

@@

%\fxnote{Dati neku smislenu procjenu kompleksnosti algoritma, ako to nije pretesko. Aca: zasad bih ovo izbegao, nezgodno je. Ako budu trazili recenzenti, uradicemo.}

@@

Method named \texttt{Replace}$(s, n, nc)$ inside Algorithm~\ref{alg:pertNode} is simply used to replace the node $n$ with node $nc$ inside the given expression tree $s$.    
%\fxnote{Nekako zbunjuje u Algoritmu 4, pojava NodeChanges)() i Replace(). Mislim, razumijem iz opisa sta one rade, ali nisam siguran da treba imenovati metode, ako ih barem u kontekstu opisa ne pomenemo direktno po nazivu. ACA: za NodeChanges stoji opis operacija iznad, sad sam ubacio i u tekst da je to sustina metode NodeChanges. Za replace sam takodje dao objasnjenje.}

%\fxnote{Dati kompleksnost velicine skupa perturbacija u odnosu na velicinu drveta i dopustene promjene. ACA: tesko je ovo izracunati, u principu za svaki cvor, dakle, celo drvo ili bilo koje pod-drvo ide ovaj spisak gore, dakle, neki red velicine d*C(T), gde je C(T) broj poddrveca iz datog drveta T, a d je neka konstanta, tj. broj mogucih perturbacija jednog poddrveta. Sad koliko ima poddrveca, moze se izracunati ovako nekako, ali ne bih to ukljucivao osim ako nam ne zatraze. Dakle, za drvo T broj poddrveca $C(T) = C(T.left)+C(T.right)+1$ (ovo jedan za bas to celo drvo u datom rekurzivnom koraku), sad kad je drvo T od N cvorova balansirano, dobija se diferencna jednacina $C(N)=2*C(N/2)+1$, a njeno resenje je O(N) mislim. Kad se raspise dobija se $C(N) = 2*2*C(N/4)+2 = 2*2*2*C(N/8)+3 = ... = 2^(\log_2(N))*C(1)+\log_2(N) = N + \log_2(N) = O(N)$ posto je $C(1)=1$, to je list drveta}

@@

For a given perturbation $p$, local search systematically explores extended set of 1-perturbations around $p$. It relies on the \emph{best-improvement} strategy, meaning that all 1-perturbations (for all subtrees) are considered. Before checking if the candidate solution ($new$) is better than the actual best $bp$, the OLS coefficient fitting (\texttt{FitOLS}) takes place.  
The set of used 1-perturbations is increased in comparison to those used previously. Namely, in addition to those six possible types of node changes, the following four are added:

@@

\fxnote{ACA: do ovde sam pregledao}
The above-mentioned benchmark sets may be biased since the models which describe physical laws usually impose   various symmetries, periodicity w.r.t. some variables, internal separability on some variables etc. In that way, the search process and the search space can be pre-defined and thus significantly reduced. Thereby, some methods may get significant benefit with this regard, while some others may be in a subordinate situation. 
In order to permit any kind of favoritism caused by physical lows behind the input data, we generate a set of randomly generated (non-biased) SR problem instances; this benchmark set is therefore labelled by \textsc{Random}. These instances are generated according the following procedure. \fxnote{Opisati proceduru generisanja instanci + kombinaciju parametara za instance + koliko instanci, itd.}

Another purpose of having these instances is  systematically  providing conclusions on robustness and scalability of the RILS-ROLS method w.r.t.\ various instance size parameters.

@@

Each table consists of three blocks. The performances of all (14) algorithms are displayed in respective rows.  
The first block gives the label of the algorithm  whose performance is reported. The second block reports by two columns the relative percentages of all (10) algorithm runs on all problem instances for which the exact model has been found by respective algorithm  on \textsc{Feynman} and \textsc{Strogatz} benchmark set, respectively. The third column of this block provides the overall   relative exact percentage of success of the algorithm over all runs on all problem instances from the both benchmark sets. The third block displays the results of the respective algorithm in terms of   final $R^2$ score which indicate  on accuracy of found solution. The solution is considered accurate if it     produced an $R^2$ score larger that  0.999.  Note that the termination of our algorithm may not always be triggered by reaching an appropriate $R^2$ score, but also due to the reached time limit. These kind of results are of a particular interest as they may point out on the presence of strong over-fitting of algorithms in contrast to the primary goal, that is finding the right model. This block consists of three columns that provide the relative percentage of algorithm runs on the instances where he algorithm is able to reach the desired accuracy (fulfilling the condition $R^2 > 0.999$ upon termination) on the both ground--truth benchmark sets separately, and the overall percentage ov all (10) runs on all considered problem instances.   %) w.r.t. the criterion of reaching the final accuracy of 0.999.

%Note that when the algorithm terminates with $R^2=1$ and $RMSE=0$, the obtained model and the exact model are always match (but also checked manually). However, in case when the algorithms deliver an error $R^2 \geq 0.999$, but $R^2 \neq 1$, these cases are all  manually checked by comparing weather the produced model match the exact model.

@@

%time-avg-noise0.0.eps

@@

	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[  width=220pt,height=150pt]{plots/time-avg-noise0.0.eps}
		\caption{No noise}
		\label{fig:noNoise-time}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.45\textwidth}

		\includegraphics[  width=220pt,height=150pt]{plots/time-avg-noise0.001.eps}
		\caption{Level of noise equal to 0.001}
		\label{fig:noise0.001-time}
	\end{subfigure}
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[  width=220pt,height=150pt]{plots/time-avg-noise0.01.eps}
		\caption{Level of noise equal to 0.01.}
		\label{fig:noise0.01-time}
	\end{subfigure}
	\caption{Average runtimes on the instance problems where the exact solution has  been reached (considered only those algorithms whose the exact percentage is at least 5\% of all runs).}
	\label{fig:time-avg-exact-noise}
\end{figure}

@@

\begin{figure}[!ht]
	 	\centering
	\begin{subfigure}[b]{0.6\textwidth} 
		\includegraphics[  width=340pt,height=180pt]{plots/Feynman_strogatz_noise_0_0-expanded.pdf}
		\caption{No noise}
		\label{fig:CDplots-no-noise}
	\end{subfigure}

	\begin{subfigure}[b]{0.6\textwidth}

		\includegraphics[  width=340pt,height=180pt]{plots/Feynman_strogatz_noise_0_001-expanded.pdf}
		\caption{Level of noise equal to 0.001}
		\label{fig:CDplots-noise0.001}
	\end{subfigure}

	\begin{subfigure}[b]{0.6\textwidth}
		\includegraphics[  width=340pt,height=180pt]{plots/Feynman_strogatz_noise_0_01-expanded.pdf}
		\caption{Level of noise equal to 0.01. }
		\label{fig:CDplots-noise0.01}
	\end{subfigure}
	\caption{CD plots comparisons in terms of exact percentages over all problem instances from the both ground-truth benchmark sets,  varying levels of noise.  }
	\label{fig:cd-plots-exact-pcts}
\end{figure}

@@

\begin{figure}[H]
	\begin{subfigure}[b]{0.40\textwidth}
		\includegraphics[  width=180pt,height=140pt]{plots/numsize_vs_exact_correct_no_noise.eps}
		\caption{No noise}
		\label{fig:size-no-noise}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.4\textwidth}

@@

\begin{figure}[H]
	\begin{subfigure}[b]{0.40\textwidth}
		\includegraphics[  width=180pt,height=140pt]{plots/numsize_vs_time_no_noise.eps}
		\caption{No noise}
		\label{fig:time-no-noise}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.4\textwidth}

@@

\begin{comment}
	\begin{center}
		\begin{tikzpicture}
			\begin{axis}[
				xlabel=$variable\;count\;in\;exact\;models$,
				ylabel=$exact\;percentage$,
				xmin=0, xmax=5,
				ymin=0, ymax=100,
				xtick={1,2,3,4},
				xticklabels={1,2,3,4},   % <---
				ytick={0,10,...,100}
				]
				\addplot[smooth,mark=*,blue] plot coordinates {
					(1,43.33)
					(2,43.33)
					(3,46.67)
					(4,46.67)
				};
				\addlegendentry{No noise}

				\addplot[smooth,color=red,mark=x]
				plot coordinates {
					(1,3.33)
					(2,3.33)
					(3,26.67)
					(4,26.67)
				};
				\addlegendentry{Noise level 0.001}

				\addplot[smooth,color=green,mark=o]
				plot coordinates {
					(1,3.33)
					(2,3.33)
					(3,3.33)
					(4,16.67)
				};
				\addlegendentry{Noise level 0.01}
			\end{axis}
		\end{tikzpicture}
		\captionof{figure}{Exact solution percentages for varying levels of noise and variable counts (only sizes ranging from 7 to 12)}
		\label{fig:compExact_noise_varcnt}
	\end{center}
\end{comment}
%\begin{center}
%	\begin{tikzpicture}
	%		\begin{axis}[
		%			xlabel=$variable\;count$,
		%			ylabel=$exact\;percentage$,
		%			xmin=0, xmax=5,
		%			ymin=0, ymax=100,
		%			xtick={1,2,3,4},
		%			xticklabels={1,2,3,4},   % <---
		%			ytick={0,10,...,100}
		%			]
		%			\addplot[smooth,mark=*,blue] plot coordinates {
			%				(1,83.33)
			%				(2,66.67)
			%				(3,73.33)
			%				(4,46.67)
			%			};
		%			\addlegendentry{No noise}
		%			
		%			\addplot[smooth,color=red,mark=x]
		%			plot coordinates {
			%				(1,56.67)
			%				(2,63.33)
			%				(3,66.67)
			%				(4,46.67)
			%			};
		%			\addlegendentry{Noise level 0.001}
		%			
		%			\addplot[smooth,color=green,mark=o]
		%			plot coordinates {
			%				(1,23.33)
			%				(2,43.33)
			%				(3,53.33)
			%				(4,26.67)
			%			};
		%			\addlegendentry{Noise level 0.01}
		%		\end{axis}
	%	\end{tikzpicture}
%	\captionof{figure}{Percentages of solutions having $R^2 > 0.999$ for varying levels of noise and variable counts (only sizes ranging from 7 to 12)}
%	\label{fig:compR2_noise_varcnt}
%\end{center}

@@

\newpage
%\section*{References}
\bibliographystyle{abbrv}	
\bibliography{bib}

\newpage
\appendix

\section{Overview of \textsc{RILS}-\textsc{ROLS} python package}\label{sec:appendix-1}

@@

%\section{Additional statistical analysis}\label{sec:appendix-2}

>>>>

%% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognise this template!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%          <8 June 2012>              %%
%%                                     %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.html and the instructions for   %%
%% authors page on the biomed central website                      %%
%% https://www.biomedcentral.com/getpublished                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% https://miktex.org/                                             %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

%%% loading packages, author definitions

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass{bmcart}

%%% Load packages
\usepackage{amsthm,amsmath}
\usepackage{tabularx}
\usepackage{xspace}
\usepackage{color}
\usepackage{epsfig}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{mathrsfs}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}

\usepackage{tikz}
%\usepackage{fullpage}
\usepackage{calc}
\usetikzlibrary{positioning,shadows,arrows,trees,shapes,fit}
\usepackage{blindtext}
\usepackage{pgfplots}
%\pgfplotsset{compat=1.16}
\usepackage{pythonhighlight}

\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{url}
\raggedbottom
%\algnewcommand\algorithmicforeach{\textbf{for each}}
%\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}

@@

\usepackage{fixme}

%\RequirePackage[numbers]{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
%\RequirePackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\def\includegraphic{}
%\def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
\endlocaldefs

%%% Begin ...
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{ {RILS}-{ROLS} : Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author[
  addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
  corref={aff1},                       % id of corresponding address, if any
% noteref={n1},                        % id's of article notes, if any
  email={kartelj@matf.bg.ac.rs}   % email address
]{\inits{A.K.}\fnm{Aleksandar} \snm{Kartelj}}
\author[
  addressref={aff2},
  email={marko.djukanovic@pmf.unibl.org}
]{\inits{M.DJ.}\fnm{Marko} \snm{Djukanović}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address[id=aff1]{%                           % unique id
  \orgdiv{Department of Informatics, Faculty of Mathematics},             % department, if any
  \orgname{University of Belgrade},          % university, etc
  \city{Belgrade},                              % city
  \cny{Serbia}                                    % country
}
\address[id=aff2]{%
  \orgdiv{Faculty of Sciences and Mathematics},
  \orgname{University of Banja Luka},
  %\street{},
  %\postcode{}
  \city{Banja Luka},
  \cny{Bosnia and Herzegovina}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter short notes here                   %%
%%                                          %%
%% Short notes will be after addresses      %%
%% on first page.                           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{artnotes}
%%\note{Sample of title note}     % note to the article
%\note[id=n1]{Equal contributor} % note, connected to author
%\end{artnotes}

\end{fmbox}% comment this for two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                           %%
%% The Abstract begins here                  %%
%%                                           %%
%% Please refer to the Instructions for      %%
%% authors on https://www.biomedcentral.com/ %%
%% and include the section headings          %%
%% accordingly for your article type.        %%
%%                                           %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}

\begin{abstract} % abstract
In this paper, we solve the well-known symbolic regression problem. This problem has been intensively studied in the literature and has a wide range of applications. The symbolic regression models could provide insights towards establishing physical theory behind experimentally observed data. Practical applications include revealing complex ecological dynamics in ecology, discovering mutations effects on protein stability in biology, modeling analytic representations of the exciton binding energy in energy science, finding models for band gaps of NaCl-type compounds
in material science, just to name a few.

@@

The experiments are conducted on the two well-known ground-truth benchmark sets called \textsc{Feynman} and \textsc{Strogatz}. \textsc{RILS}-\textsc{ROLS} was compared to 14 other competitors from the literature with  presence of different levels of the Gaussian white noise in the input data. Our method outperformed all 14 competitors with respect to the true symbolic model accuracy percentage under all level of noise. 
In addition to evaluation on known ground-truth datasets, a new randomly generated set of problem instances is introduced. The goal of \textsc{Random} dataset was to test the scalability of our method with respect to incremental equation sizes and different levels of noise. \textsc{RILS}-\textsc{ROLS} can be considered as a new state-of-the-art method for solving the problem of symbolic regression when ground-truth equations are known.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
\kwd{Symbolic regression}
\kwd{Iterated local search}
\kwd{Ordinary least square}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
%
%\end{fmbox}% uncomment this for two column layout

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                            %%
%% The Main Body begins here                  %%
%%                                            %%
%% Please refer to the instructions for       %%
%% authors on:                                %%
%% https://www.biomedcentral.com/getpublished %%
%% and include the section headings           %%
%% accordingly for your article type.         %%
%%                                            %%
%% See the Results and Discussion section     %%
%% for details on how to create sub-sections  %%
%%                                            %%
%% use \cite{...} to cite references          %%
%%  \cite{koon} and                           %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}      %%
%%                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
% <put your article body there>

%%%%%%%%%%%%%%%%
%% Background %%
%%
	\section{Introduction}\label{sec:introduction}

@@

@@

Finally, by setting each partial derivative to zero, we get the system of $d+1$ linear equations with $d+1$ unknowns ($\mathbf{k}=(k_1, ..., k_d)$ and $b$): 
 \begin{align*}
	& \sum_{i=1}^{n} -2x_{i1} (y_i - \tilde{y}_i) = 0 \\
	& \vdots \\
	& \sum_{i=1}^{n} -2x_{id} (y_i - \tilde{y}_i) = 0 \\
	&\sum_{i=1}^{n} -2  (y_i - \tilde{y}_i) = 0.
\end{align*}

@@

@@

Method named \texttt{Replace}$(s, n, nc)$ inside Algorithm~\ref{alg:pertNode} is simply used to replace the node $n$ with node $nc$ inside the given expression tree $s$.

@@

For a given perturbation $p$, local search systematically explores extended set of 1-perturbations around $p$. It relies on the \emph{best-improvement} strategy, meaning that all 1-perturbations (for all subtrees) are considered. Before checking if the candidate solution ($new$) is better than the actual best $bp$, the OLS coefficient fitting (\texttt{FitOLS}) takes place.

The set of used 1-perturbations is increased in comparison to those used previously. Namely, in addition to those six possible types of node changes, the following four are added:

@@

The above-mentioned benchmark sets may be biased since the models which describe physical laws usually impose   various symmetries, periodicity w.r.t. some variables, internal separability on some variables, etc. In order to test \textsc{RILS-ROLS} robustness and scalability in unbiased setting, we generated a set of random SR problem instances called \textsc{Random}. These instances have a varying size (total number of expression nodes) and number of variables. In total, there are 235 random instances, or 5 randomly generated instances for each of the following 47 (size, number of variables) combinations $\{(3, 1), \{4, 5\} \times \{1, 2\}, \{6\} \times \{1, 2, 3\}, \{7, 8, 9, 10, 11, 12\} \times \{1, 2, 3, 4\}, \{13, 14, 15\} \times \{1, 2, 3, 4, 5\}\}$. Each of \textsc{Random} instances has 10 thousands randomly generated samples.

Some \textsc{Random} instances are shown in Table~\ref{tbl:random}, while all instances can be found at \textsc{RILS-ROLS} GitHub repository: \url{https://github.com/kartelj/rils-rols}.

\begin{table}
	\centering
	\begin{tabular}{ll} \\ \hline
		random\_04\_02\_0010000\_00 &	$\ln{(x_0 + x_1)}$\\
		random\_07\_02\_0010000\_00 &	$(x_0 + 10)\cdot (x_1 + 1)$\\
		random\_08\_02\_0010000\_00 &	$(x_0 + \ln{x_0})*(x_1 + 10)$\\
		random\_10\_03\_0010000\_02 &	$\ln{(\sqrt{x_0} \cdot x_1/\cos{x_2})}$\\
		random\_12\_04\_0010000\_04 &	$\sqrt{x_3 \cdot \exp{x_1} + \cos{(x_2 \cdot \sin{x_0})}}$\\
		random\_15\_03\_0010000\_04	&   $x_0 \cdot x_1/2 + (\cos{(x_2)} - \pi^2)^2$\\
		\hline
	\end{tabular}
	\caption{Some \textsc{Random} instance problems.}
	\label{table:random}
\end{table}

@@

Each table consists of three blocks. The performances of all (14) algorithms are displayed in respective rows.  
The first block gives the label of the algorithm  whose performance is reported. The second block reports by two columns the relative percentages of all (10) algorithm runs on all problem instances for which the exact model has been found by respective algorithm  on \textsc{Feynman} and \textsc{Strogatz} benchmark set, respectively. The third column of this block provides the overall   relative exact percentage of success of the algorithm over all runs on all problem instances from the both benchmark sets. The third block displays the results of the respective algorithm in terms of   final $R^2$ score which indicate  on accuracy of found solution. The solution is considered accurate if it     produced an $R^2$ score larger that  0.999.  Note that the termination of our algorithm may not always be triggered by reaching an appropriate $R^2$ score, but also due to the reached time limit. These kind of results are of a particular interest as they may point out on the presence of strong over-fitting of algorithms in contrast to the primary goal, that is finding the right model. This block consists of three columns that provide the relative percentage of algorithm runs on the instances where he algorithm is able to reach the desired accuracy (fulfilling the condition $R^2 > 0.999$ upon termination) on the both ground--truth benchmark sets separately, and the overall percentage ov all (10) runs on all considered problem instances.

@@

@@

	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=190pt,height=150pt]{plots/time-avg-noise0.0.eps}
		\caption{No noise}
		\label{fig:noNoise-time}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.45\textwidth}

		\includegraphics[  width=190pt,height=150pt]{plots/time-avg-noise0.001.eps}
		\caption{Level of noise equal to 0.001}
		\label{fig:noise0.001-time}
	\end{subfigure}
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[  width=190pt,height=150pt]{plots/time-avg-noise0.01.eps}
		\caption{Level of noise equal to 0.01.}
		\label{fig:noise0.01-time}
	\end{subfigure}
	\caption{Average runtimes on the instance problems where the exact solution has  been reached (considered only those algorithms whose the exact percentage is at least 5\% of all runs).}
	\label{fig:time-avg-exact-noise}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\begin{subfigure}[b]{0.75\textwidth} 
		\includegraphics[  width=300pt,height=180pt]{plots/Feynman_strogatz_noise_0_0-expanded.pdf}
		\caption{No noise}
		\label{fig:CDplots-no-noise}
	\end{subfigure}

	\begin{subfigure}[b]{0.75\textwidth}

		\includegraphics[  width=300pt,height=180pt]{plots/Feynman_strogatz_noise_0_001-expanded.pdf}
		\caption{Level of noise equal to 0.001}
		\label{fig:CDplots-noise0.001}
	\end{subfigure}

	\begin{subfigure}[b]{0.75\textwidth}
		\includegraphics[  width=300pt,height=180pt]{plots/Feynman_strogatz_noise_0_01-expanded.pdf}
		\caption{Level of noise equal to 0.01. }
		\label{fig:CDplots-noise0.01}
	\end{subfigure}
	\caption{CD plots comparisons in terms of exact percentages over all problem instances from the both ground-truth benchmark sets,  varying levels of noise.  }
	\label{fig:cd-plots-exact-pcts}
\end{figure}

@@

\begin{figure}[!ht]
	\begin{subfigure}[b]{0.40\textwidth}
		\includegraphics[  width=180pt,height=140pt]{plots/numsize_vs_exact_correct_no_noise.eps}
		\caption{No noise}
		\label{fig:size-no-noise}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.4\textwidth}

@@

\begin{figure}[!ht]
	\begin{subfigure}[b]{0.40\textwidth}
		\includegraphics[  width=180pt,height=140pt]{plots/numsize_vs_time_no_noise.eps}
		\caption{No noise}
		\label{fig:time-no-noise}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.4\textwidth}

@@

@@

\section*{Appendix}
\section{Overview of \textsc{RILS}-\textsc{ROLS} python package}\label{sec:appendix-1}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

\section*{Acknowledgements}%% if any
Text for this section\ldots

\section*{Funding}%% if any
Text for this section\ldots

\section*{Abbreviations}%% if any
Text for this section\ldots

\section*{Availability of data and materials}%% if any
Text for this section\ldots

\section*{Ethics approval and consent to participate}%% if any
Text for this section\ldots

\section*{Competing interests}
The authors declare that they have no competing interests.

\section*{Consent for publication}%% if any
Text for this section\ldots

\section*{Authors' contributions}
Text for this section \ldots

\section*{Authors' information}%% if any
Text for this section\ldots
